Equilibria and Convergence of Fictitious Play\\ on Network Aggregative Games

### 2021-05-28 Documents/FP in NAGs/neurips_2021.tex
Updated according to comments

\paragraph{Contributions.}
%
  The main contribution of this work is to introduce and study convergence of
  learning on Network Aggregative games through the action of
  Fictitious Play (FP). In particular, we aim to establish guarantees
  of convergence for learning under certain classes of games. Such guarantees
  allow for games to be designed so that the long-term behaviour of
  learning is predictable, regardless of how many agents are engaged
  in the game. As such, we are able to avoid the pitfalls of chaotic
  dynamics in games with many players, as shownin \cite{Sanders2018}.

@@

\begin{remark}
	%It can be seen that
	The notion of Nash equilibrium in NA Games is a natural
	extension of the NE in a bimatrix game.
	%$(A, B)$.
	In particular, if we consider an NA game with only two players and no
	self-loops then the above definition yields that $ \NE{1}$ is an NE iff
	%
	%    \begin{equation*}
	$     \NE{1} \in BR^1 (\sigma^1) = \arg \max_{x in \Delta_1} u^1 (x, \NE{2})$,
	%    \end{equation*}
	%
	and similarly for $\NE{2}$. This is precisely the definition of an NE in a two player game \cite{}.
\end{remark}

@@

  \begin{remark}
    As an illustration, consider NA Games with two players, in which $\edgeset = \{(1, 2),
    (2, 1)\}$ and $\weightset$ is a 2x2 matrix with zeros on its leading diagonal and ones on
    the off diagonal. We write the time-average of both agents' state as
    \begin{align*}
      \alpha^\mu(t; x) = \frac{1}{t} \int_0^t \xmu(s) \, ds & \text{ for $\mu \in \{1, 2\}$}
  %   \alpha^2 = \frac{1}{t} \int_0^T x^2(t) \, dt \\
    \end{align*}
    In this manner, the $\alpha^\mu(t; m)$ denotes the time average of the strategies played by
    agent $\mu$ up to time $t$ when the strategies are given by $\xmu(t)$. Note that we often reduce the notation to $\alpha^\mu(t)$. Then, fictitious play requires that the agents update their
    strategy as $x^1(t) \in BR^1(\alpha^2(t))$ and $x^2(t) \in BR^2(\alpha^1(t))$. It can be
    seen, therefore, that the NA-CTFP is a natural extension of CTFP in the classical
    two-player setting \cite{Hofbauer2006}.
  \end{remark}

@@

  \begin{remark}
    As usual, the notion of CCE in an NA game is a natural extension of the CCE for two player
    games. In fact, if we consider the NA game to be a two player game with no self-loops, then
    we recover exactly the definition of the CCE set in two player games \cite{Ostrovski2014}.
  \end{remark}

  \begin{remark} \label{rem::regretvsCCE}
    The notion of the CCE set is related to the idea of \emph{average regret}  \cite{Nisan2007}. Here, we
    will present what is meant by average regret and state that if at some time $t$ all agents'
    average regret is non-positive, then the game is said to have reached the CCE set. The
    reader should consult \cite{Ostrovski2014} for an excellent exposition regarding the
    link between the CCE set and average regret in two player games which, of course, extends
    naturally to NA Games.

@@

    in which $\pure{\mu}{i}$ denotes the probability vector in $\Delta_\mu$ with $1$ in the slot
    $i'$ and $0$ everywhere else. Note, this is the \emph{average regret} for the agent $\mu$
    and, of course, can be related to the \emph{cumulative regret} which is used for analysis
    in \cite{Leonardos2020, Cesa-Bianchi2021}.  To illustrate the average regret, let us
    consider the case where each agent has only two actions. Then $u^{\mu}(x^\mu(t), \sigma(t))$
    is given by
%    
    \begin{equation} \label{eq1}
      u^{\mu}(x^\mu(t), \sigma(t)) = \sum_{ij} a_{ij} x_i^\mu \sigma_j^\mu = a_{11} x_1^\mu \sigma_1^\mu + a_{12} x_1^\mu \sigma_2^\mu + a_{21} x_2^\mu \sigma_1^\mu + a_{22} x_2^\mu \sigma_2^\mu
    \end{equation}

@@

  As an adaptation, we take the example of a three player chain, in which player $2$ is connected to $1$ and $3$. The aggregation matrix can be given as
%
  \begin{equation*}
    W = \begin{bmatrix}
      0 & 1 & 0 \\
      w & 0 & 1 - w \\
      0 & 1 & 0
    \end{bmatrix}, \; w \in (0, 1).
  \end{equation*}

  We first consider the zero-sum case to show that it does indeed converge to an equilibrium as expected. Note that the zero-sum condition (Assumption \ref{ass::zerosum}) given for the three player chain is given as
%
  \begin{equation}
    x \cdot A y + y \cdot B (w x + (1-w)z) + z \cdot C y = 0. \hspace{0.5cm} \forall x, y, z \in \Delta_1 \times \Delta_2 \times \Delta_3
  \end{equation}
  %
  in which we use the notation that $x, y, z$ (resp. $A, B, C$) denote the strategies (resp. payoffs) of agents 1, 2 and 3 respectively. This condition is satisfied if we fix $B$ and choose
%
  \begin{align} \label{eq::zeroSumShapley}
    A & = - w B^T \nonumber \\
    C & = - (1 - w) B^T. 
  \end{align}

@@

  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8\textwidth]{Figures/convergentShapley.png}
    \caption{\label{fig::convergentShapley}}
    \end{subfigure}
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8 \textwidth]{Figures/nonConvergentShapley.png}
      \caption{\label{fig::nonconvergentShapley}}
    \end{subfigure}
    \caption{\label{fig::Shapley} Orbits of the Fictitious Play in the Three Player Chain for which the NE lies at the centre of the simplex. (a) Payoffs are given by
    (\ref{eq::zeroSumShapley}). The plot shows NA-CTFP yields convergence to the NE (b)
	NA-CTFP showing cycles around the NE when payoffs are modified slightly (c.f. Sec.~\ref{sec::NonConv}).}
  \end{figure}

@@

  \begin{figure}[t]
    \centering
    \includegraphics[width = 0.9 \columnwidth]{Figures/Noise20Player.png}
    \caption{\label{fig::Noise20Player} Trajectories of NA-CTFP in a 20 player game with additive noise (Left) No noise is introduced and learning converges directly to an NE. (Middle) $\gamma = 0.05$, the trajectories converge to a fixed point but removed from the NE. (Right) $\gamma = 0.1$, the trajectories converge to a fixed point which is even further away from the NE.}
  \end{figure}

@@

  The fictitious play process in NA games requires that, at each time step, an agent takes a
  `measurement' of the aggregate strategy of its neighbours. It is on this measurement that they
  update their own strategy. It stands to reason then, that in real environments this measurement
  may be corrupted by noise. 
%  
  As such, we investigate the effect that introducing additive noise has
  on NA-CTFP in a zero-sum NA game. We do this in the following manner: at each time step, the
  reference signal $\refmu(t)$ is adjusted to $\refmu + \gamma \xi$ where $\xi$ is drawn from the
  standard normal distribution (zero mean and unit variance). By varying $\gamma$, we vary the
  strength of the noise. We vary $\gamma$ up to $0.5$ since, above this value, noisy measurements
  are likely to lie outside of the simplex. Since $\refmu$ is constrained to lie within $\Delta$, we
  can consider the range $\gamma \in [0, 0.5]$ to be the \emph{physical region}, in which noise is meaningful.

@@

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit. However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Smith1982}, another adaptive algorithm linked to multi-agent learning
  \cite{Mertikopoulos2018}. In \cite{Imhof2005} and \cite{Galla2011}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.

@@

	Our work opens a number of lines for future work. Most notable is the effect of noise. It would be prudent to analyse this theoretically, as was done in \cite{Perrin2020}, and consider the conditions under which FP will still converge to a fixed point. Furthermore, it would be interesting to investigate the phenomenon we report experimentally in a theoretical framework. Namely, the question of why noise breaks periodicity in FP and results in convergence to an NE should be investigated and, indeed, this is a line which we are currently pursuing.
%	
	In addition, we note that the \emph{Mann Iteration}, a method of approximating fixed points which is investigated in \cite{Parise2020}, shares a remarkably similar structure to the discrete variant of FP. This may present an avenue by which NA-CTFP may be analysed in the case of convex cost functions. 
%	
	Finally, we note that in recent years FP in two player games has shown a remarkable variety of dynamical behaviours, including periodicity and chaos. In our work we have shown convergence to a fixed point and, through experiments, periodicity. It stands to reason, therefore, that a greater variety of dynamical behaviours exist for NA-CTFP for certain classes of games. It would be important to determine what these classes are. Short from being merely a curiosity, this would allow for the identification of games in which NA-CTFP leads to inherently unpredictable behaviour, an important question from the point of view of building Safe and Trusted AI.

>>>>

\paragraph{Contributions.}
%
  The main contribution of this work is to introduce and study convergence of
  learning on Network Aggregative games through the action of
  Fictitious Play (FP). In particular, we aim to establish guarantees
  of convergence for learning under certain classes of games. Such guarantees
  allow for games to be designed so that the long-term behaviour of
  learning is predictable, regardless of how many agents are engaged
  in the game. As such, we are able to avoid the pitfalls of chaotic
  dynamics in games with many players (also called `agents'), as shownin \cite{Sanders2018}.

@@

\begin{remark}
	%It can be seen that
	The notion of Nash equilibrium in NA Games is a natural
	extension of the NE in a bimatrix game.
	%$(A, B)$.
	In particular, if we consider an NA game with only two players and no
	self-loops then the above definition yields that $ \NE{1}$ is an NE iff
	%
	%    \begin{equation*}
	$     \NE{1} \in BR^1 (\sigma^1) = \arg \max_{x in \Delta_1} u^1 (x, \NE{2})$,
	%    \end{equation*}
	%
	and similarly for $\NE{2}$. This is precisely the definition of an NE in a two-player game \cite{}.
\end{remark}

@@

  \begin{remark}
    As an illustration, consider NA Games with two players, in which $\edgeset = \{(1, 2),
    (2, 1)\}$ and $\weightset$ is a 2x2 matrix with zeros on its leading diagonal and ones on
    the off diagonal. We write the time-average of both agents' state as
    \begin{align*}
      \alpha^\mu(t; x) = \frac{1}{t} \int_0^t \xmu(s) \, ds & \text{ for $\mu \in \{1, 2\}$}
  %   \alpha^2 = \frac{1}{t} \int_0^T x^2(t) \, dt \\
    \end{align*}
    In this manner, the $\alpha^\mu(t; m)$ denotes the time average of the strategies played by
    agent $\mu$ up to time $t$ when the strategies are given by $\xmu(t)$. Note that we often reduce the notation to $\alpha^\mu(t)$. Then, Fictitious Play requires that the agents update their
    strategy as $x^1(t) \in BR^1(\alpha^2(t))$ and $x^2(t) \in BR^2(\alpha^1(t))$. It can be
    seen, therefore, that the NA-CTFP is a natural extension of CTFP in the classical
    two-player setting \cite{Hofbauer2006}.
  \end{remark}

@@

  \begin{remark}
    As usual, the notion of CCE in an NA game is a natural extension of the CCE for two-player
    games. In fact, if we consider the NA game to be a two-player game with no self-loops, then
    we recover exactly the definition of the CCE set in two-player games \cite{Ostrovski2014}.
  \end{remark}

  \begin{remark} \label{rem::regretvsCCE}
    The notion of the CCE set is related to the idea of \emph{average regret}  \cite{Nisan2007}. Here, we
    will present what is meant by average regret and state that if at some time $t$ all agents'
    average regret is non-positive, then the game is said to have reached the CCE set. The
    reader should consult \cite{Ostrovski2014} for an excellent exposition regarding the
    link between the CCE set and average regret in two-player games which, of course, extends
    naturally to NA Games.

@@

    in which $\pure{\mu}{i}$ denotes the probability vector in $\Delta_\mu$ with $1$ in the slot
    $i'$ and $0$ everywhere else. Note, this is the \emph{average regret} for the agent $\mu$
    and, of course, can be related to the \emph{cumulative regret} which is used for analysis
    in \cite{Cesa-Bianchi2021}.  To illustrate the average regret, let us
    consider the case where each agent has only two actions. Then $u^{\mu}(x^\mu(t), \sigma(t))$
    is given by
%    
    \begin{equation} \label{eq1}
      u^{\mu}(x^\mu(t), \sigma(t)) = \sum_{ij} a_{ij} x_i^\mu \sigma_j^\mu = a_{11} x_1^\mu \sigma_1^\mu + a_{12} x_1^\mu \sigma_2^\mu + a_{21} x_2^\mu \sigma_1^\mu + a_{22} x_2^\mu \sigma_2^\mu
    \end{equation}

@@

  As an adaptation, we take the example of a three-player chain, in which player $2$ is connected to $1$ and $3$. The aggregation matrix can be given as
%
  \begin{equation*}
    W = \begin{bmatrix}
      0 & 1 & 0 \\
      w & 0 & 1 - w \\
      0 & 1 & 0
    \end{bmatrix}, \; w \in (0, 1).
  \end{equation*}

  We first consider the zero-sum case to show that it does indeed converge to an equilibrium as expected. Note that the zero-sum condition (Assumption \ref{ass::zerosum}) given for the three-player chain is given as
%
  \begin{equation}
    x \cdot A y + y \cdot B (w x + (1-w)z) + z \cdot C y = 0. \hspace{0.5cm} \forall x, y, z \in \Delta_1 \times \Delta_2 \times \Delta_3
  \end{equation}
  %
  in which we use the notation that $x, y, z$ (resp. $A, B, C$) denote the strategies (resp. payoffs) of agents 1, 2 and 3 respectively. This condition is satisfied if we fix $B$ and choose
%
  \begin{align} \label{eq::zeroSumShapley}
    A & = - w B^T \nonumber \\
    C & = - (1 - w) B^T. 
  \end{align}

@@

  \begin{figure}[t]
    \centering
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8\textwidth]{Figures/convergentShapley.png}
    \caption{\label{fig::convergentShapley}}
    \end{subfigure}
    \begin{subfigure}[b]{0.45 \textwidth}
      \includegraphics[width = 0.8 \textwidth]{Figures/nonConvergentShapley.png}
      \caption{\label{fig::nonconvergentShapley}}
    \end{subfigure}
    \caption{\label{fig::Shapley} Orbits of the Fictitious Play in the Three-Player Chain for which the NE lies at the centre of the simplex. (a) Payoffs are given by
    (\ref{eq::zeroSumShapley}). The plot shows NA-CTFP yields convergence to the NE (b)
	NA-CTFP showing cycles around the NE when payoffs are modified slightly (c.f. Sec.~\ref{sec::NonConv}).}
  \end{figure}

@@

  \begin{figure}[t]
    \centering
    \includegraphics[width = 0.9 \columnwidth]{Figures/Noise20Player.png}
    \caption{\label{fig::Noise20Player} Trajectories of NA-CTFP in a 20-player game with additive noise (Left) No noise is introduced and learning converges directly to an NE. (Middle) $\gamma = 0.05$, the trajectories converge to a fixed point but removed from the NE. (Right) $\gamma = 0.1$, the trajectories converge to a fixed point which is even further away from the NE.}
  \end{figure}

@@

  The Fictitious Play process in NA games requires that, at each time step, an agent takes a
  `measurement' of the aggregate strategy of its neighbours. It is on this measurement that they
  update their own strategy. It stands to reason then, that in real environments this measurement
  may be corrupted by noise. 
%  
  As such, we investigate the effect that introducing additive noise has
  on NA-CTFP in a zero-sum NA game. We do this in the following manner: at each time step, the
  reference signal $\refmu(t)$ is adjusted to $\refmu + \gamma \xi$ where $\xi$ is drawn from the
  standard normal distribution (zero mean and unit variance). By varying $\gamma$, we vary the
  strength of the noise. We vary $\gamma$ up to $0.5$ since, above this value, noisy measurements
  are likely to lie outside of the simplex. Since $\refmu$ is constrained to lie within $\Delta$, we
  can consider the range $\gamma \in [0, 0.5]$ to be the \emph{physical region}, in which noise is meaningful.

@@

  In Figure \ref{fig::3PlayerChainNoise} we revisit the Three-Player Chain of Section
  \ref{sec::NonConv}, now under the influence of additive noise. For the sake of brevity, we only
  display the distance to the Nash Equilibrium of the first player's action, since the other agents
  behave in the same way. It can be seen that a small amount of noise has the effect of decreasing
  the size of the periodic orbit. However, as $\gamma$
  is increased to $0.5$, the algorithm seems to exhibit convergence to the NE. The implication is
  that the addition of noise may cause periodic behaviour to break and lead to the Nash Equilibrium.
  An interesting point to note is that this behaviour is in stark contrast to the replicator
  dynamic (RD) \cite{Smith1982}, another adaptive algorithm linked to multi-agent learning
  \cite{Mertikopoulos2018}. In \cite{Imhof2005} and \cite{Galla2011}, it was found that the
  introduction of random mutations can remove convergent behaviour and instead lead to periodicity.

@@

	Our work opens a number of lines for future work. Most notable is the effect of noise. It would be prudent to analyse this theoretically, as was done in \cite{Perrin2020}, and consider the conditions under which FP will still converge to a fixed point. Furthermore, it would be interesting to investigate the phenomenon we report experimentally in a theoretical framework. Namely, the question of why noise breaks periodicity in FP and results in convergence to an NE should be investigated and, indeed, this is a line which we are currently pursuing.
%	
	In addition, we note that the \emph{Mann Iteration}, a method of approximating fixed points which is investigated in \cite{Parise2020}, shares a remarkably similar structure to the discrete variant of FP. This may present an avenue by which NA-CTFP may be analysed in the case of convex cost functions. 
%	
	Finally, we note that in recent years FP in two-player games has shown a remarkable variety of dynamical behaviours, including periodicity and chaos. In our work we have shown convergence to a fixed point and, through experiments, periodicity. It stands to reason, therefore, that a greater variety of dynamical behaviours exist for NA-CTFP for certain classes of games. It would be important to determine what these classes are. Short from being merely a curiosity, this would allow for the identification of games in which NA-CTFP leads to inherently unpredictable behaviour, an important question from the point of view of building Safe and Trusted AI.
